NURM: a tool to locate numerical differences in pipelines

### 2019-10-30 paper.tex
edit images and captions

\usepackage{subcaption}
\usepackage{xspace}
\usepackage{stmaryrd} % for llbracket and rrbracket
\usepackage{amsmath}
\usepackage{ulem}
\DeclareMathOperator*{\argmin}{argmin}

@@

Recently, validation of the reproducibility has been widely investigated 
in the field of neuroimaging in which uses optimized 
processing methods with the purpose of functional and structural 
assessments of the human brain.
According to the previous 
studies, the variety of computing infrastructures including workstation 
types, parallelization methods, operating systems, and analysis 
packages are known to influence the reproducibility of the analyses~\cite{Gronenschild2012, 
diethelm2012limits, Glatard2015, bowring2019exploring}.

In particular, the studies on the effect of the operating system on the results show that 
computational pipelines create small numerical differences~\cite{Glatard2015, Scaria2017}.
These differences mainly correspond to the mathematical functions implemented 
in different operating system libraries.
For instance, changing the mathematical functions like \emph{expf()} and 
\emph{cosf()} which manipulate the precision of floating-point representations, 
between \emph{glibc} libraries in different operating systems can produce 
small numerical differences.
Besides, a similar issue is expected for any operating system which is 
based on \emph{glibc}, the GNU C library.

These irreproducibility issues are reported as the result of the 
creation, propagation, and amplification of small numerical 
differences~\cite{Gronenschild2012, diethelm2012limits, Glatard2015, 
bowring2019exploring}. In this case, the analysis pipelines are said to 
be numerically unstable. Numerical instability is a characteristic of 
the pipelines which amplify small numerical differences and then hamper 
the reproducibility of the analyses depending on the length of the 
pipeline and magnitude of the differences. In many cases, numerical 
instability is an essential issue for reproducibility.

There are different solutions to improve the reproducibility of the analysis, 
including containerization techniques that encapsulate software/hardware 
dependencies, provenance capturing tools, and version control systems. 
However, a comprehensive solution requires to fix the numerical 
instabilities instead of masking the issue. For this purpose, we introduce 
the Numerical Reproducibility Measurement (NURM) tool to identify the processes in 
the pipeline that create numerical differences across several runs. 
Furthermore, we can fix the instability of the identified processes 
in the next steps.

The remainder of this paper describes the NURM-tool in different parts 
including the way of capturing and representation of the provenance 
information, the method of clustering of different subject types, and 
process labelling to characterize differences. We describe the HCP 
preprocessing pipelines as a popular neuroimaging project to test the 
NURM-tool and evaluate its functionality. Finally, we reported the 
experimental results which describe the processes that are responsible 
for the differences in the pipelines.

@@

\begin{figure}
\centering
  \includegraphics[width=0.3\columnwidth]{images/simple_graph}
  \caption{Process tree and graph
  constructed from the example pipeline in
  Listing~\ref{listing:sample-script}.
  Processes that read or write
  temporary files are 
  represented with squares. Plain edges 
  represent the process tree (\texttt{fork()} or \texttt{clone()} 
  system calls). Dashed edges represent file dependencies: temporary 
  files are in yellow and result files are in green.
  Every node in the tree is labeled using (1) a process id created by our
  reconstruction, (2) the name of the executable run by the process.
  Process 0 is the initial call to the \texttt{sh} interpreter, and
  processes 1, 3 and 4 are the calls to FSL bet, stats and maths made in
  Listing~\ref{listing:sample-script}. Process 2 is forked by process 1: it
  was captured by \reprozip while it did not appear in
  Listing~\ref{listing:sample-script}. 
}
  \label{fig:simple_script}
\end{figure}

@@

We label in two categories the processes of the graph of each subject. First,
processes that read files that do not have differences and write files that
do not have differences are labeled \emph{transparent}
(Figure~\ref{fig:green}). Second, processes that read files that do not
have differences but write files that have differences \emph{create}
differences (Figure~\ref{fig:red}). In addition, processes that read files
that have differences and write files that also have differences are called
\emph{undetermined} (Figure~\ref{fig:yellow}). Processes are also labeled
undetermined when they read or write temporary files, or produce multi-write
files. To resolve undetermined processes, hooks must be added to the pipeline to 
execute these processes on files without differences, and to capture
intermediary file states.

\begin{figure}%\centering
\centering
    \begin{subfigure}{0.2\linewidth}
        \includegraphics[scale=0.3]{images/green.png}
        \caption{Transparent}
        \label{fig:green}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.2\linewidth}
    \includegraphics[scale=0.3]{images/red.png}
    \caption{Creates differences}
    \label{fig:red}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.2\linewidth}
    \includegraphics[scale=0.3]{images/yellow.png}
    \caption{Undetermined}
    \label{fig:yellow}
\end{subfigure}
    \caption{Different type of labeled process based on the input/output files.
  Dashed edges refer to the file dependencies between processes A and B 
  if a file written by A is read by B. Solid black edges refer to the 
  relationship between parent and child processes.}
    \label{fig:processes}
\end{figure}

@@

  # Run the original executable using the same arguments
  executed_file = sys.argv[0]
  exec_argv = op.join('original_scripts', executed_file)
  exec_argv = exec_argv + " ".join(map(pipes.quote, sys.argv[1:]))
  subprocess.Popen(exec_argv, shell=True).communicate()

@@

  if exec_argv in process_list.keys():
      # Save output files to read-only directory
      p_files = process_list[exec_argv]
      shutil.copy(p_files, '/read-only/dir')

@@

Since all the result files are captured in Condition 1, the tool 
is ready to label the piepeline processes based on the results in Condition 2.
However, another challenge is to label undetermined processes that read files with 
differences and write files with differences too. In this
situation, it is not possible to determine from the pipeline results 
whether the process created differences or 
was transparent.
To address this issue, we set hooks in the pipeline 
in Condition 2. Hooks are set by replacing the pipeline 
processes with a custom script. This custom script labels each process 
as \emph{transparent} or \emph{create differences} immediately after running 
the original executable. 
The replacement is done through the PATH variable, as before. Moreover, 
this custom script copies the results obtained from the same process in 
Condition 1 to the pipeline output in Condition 2 if 
it is invoked with the arguments of a process that created differences 
(Figure~\ref{fig:overview-labelling} - (3)). 
Differences are identified by comparing checksum of the output files for 
the same process in both conditions (Figure~\ref{fig:overview-labelling} - (2)).
In addition, the distance functions that are specific to file types can be added 
to our tool to ensure that differences belong to the file data part.
We developed an incremental approach that consists of the following steps:

@@

Furthermore, the average time per subject taken for PreFreeSurfer processing is 2.2 hours, 
and FreeSurfer processing is 8.3 hours (20 runs), 
and the average output file size of PreFreeSurfer processing is 2.7 GB, 
and FreeSurfer processing is 4.1 GB (20 subjects).

HCP data collected from different types of imaging techniques, including 
structural imaging (sMRI), functional imaging (fMRI), and diffusion imaging (dMRI).
The structural images include T1-weighted (T1w) and T2-weighted (T2w) images which 
help in the diagnosis of the brain injury.
The functional images, including task-based fMRI and resting-state scans, 
enable the measurement of functional activations within brain areas. 
Diffusion imaging is another kind of MRI technique, which measures 
the anatomical connectivity between regions.
See~\cite{van2013wu} for a discussion of these various terms.

Data from 1200 subjects 
is available in \href{https://db.humanconnectome.org}{the ConnectomDB repository}. 
In this paper, we randomly selected 20 unprocessed subjects from 
the HCP data release S500. All the subjects were scanned using the 3-tesla 
type of MRI scanners.

@@

We didn't observe any within-OS difference in PreFreeSurfer, however, we
identified two FreeSurfer processes leading to within-OS differences due
to the use of pseudo-random numbers: image registration with
\emph{mri\_segreg}, and cortical surface curvature
estimations with \emph{mris\_curvature}. To address this issue, we fixed
the seed used by the pseudo-random number generator, through an option of
the FreeSurfer tools. In this way, our analysis can focus on between-OS
differences.

@@

\begin{tabular}{@{}llllll@{}}
\toprule
Type   & Number of Subjects  & Missing File(s)                                    \\ \midrule
type1  & 9                   & None                                               \\
type2  & 7                   & [``T1w", ``T2w"]                                   \\
type3  & 2                   & [``T1w"]                                           \\
type4  & 2                   & [``T2w"]                                           \\ \bottomrule
\end{tabular}
\begin{tablenotes}
     \small
     \item *Type refers to the topology type of the process tree of subjects.
\end{tablenotes}
\end{threeparttable}
\end{table}

Figure~\ref{fig:pfs_freq} shows the frequency at which  processes create differences 
in the pipeline, and how differences are distribute among sub-pipelines.
Differences were observed in linear registration 
with FSL FLIRT (in ACPC-Alignment, Brain Extraction, Distortion Correction, 
Atlas Registration), in non-linear registration with FSL FNIRT (in Brain Extraction 
and Altas Registration), and in image warping with FSL \emph{new\_invwarp} (in Brain Extraction 
and Atlas Registration). 
Differences were also observed in image mean 
computations with FSL maths  (in Anatomical Average).

Results show that \emph{FLIRT} is the most frequent process that produces 
differences in ACPC-Alignment across all the subjects, Distortion Correction 
in 95\% of the subjects, and Atlas Registration in 85\% of the subjects. 
In addition, there is a \emph{fslmaths} process in Anatomical Average that 
creates differences in all 11 subjects in which it has appeared.

\begin{figure}
\centering
  \includegraphics[width=\columnwidth]{images/pfs_heatmap.png}
  \caption{Heatmap of difference frequency of processes along with the sub-pipelines that introduce 
           differences among 20 subjects. The range between red and green colors show the most and 
           least frequency of processes respectively. Besides, the total number 
           of occurrences of each processes is represented in the bracket in each cell.}
  \label{fig:pfs_freq}
\end{figure}

To visually represent the magnitude of variations, we created checkerboard 
images from the image results of one subject in two OSes. Figure~\ref{fig:fnirt_result} 
shows the checkerboard images from the results of ACPC-Alignment in PreFreeSurfer. 
Each rectangle includes the replaced pixels with different intensity values between images.

@@

\begin{figure}
  \centering
    \includegraphics[width=\columnwidth]{images/segmentation.png} 
    \caption{Differences between FNIRT results of one subject from PreFreeSurfer 
    (checkerboard image \emph{T2w\_acpc\_to\_MNI\_nonli.nii.gz}) (CentOS6 vs. 
    CentOS7). Right images show the zoomed version part of the images marked 
    by the red rectangle.
    Here is 
    \href{https://raw.githubusercontent.com/ali4006/HCP-reproducibility-paper/master/images/brain_classification.gif}{the link}
    to the gif animation of the same images.
} 
    \label{fig:fnirt_result}
\end{figure}

@@

In the first test, the results of PreFreeSurfer on CentOS 6
feed to the FreeSurfer pipeline on both OSes. 
We fixed input data to determine whether the pipeline creates differences.
We processed 20 subjects using FreeSurfer.
NURM-tool identified \emph{mris\_make\_surfaces} as the only process in 
the pipeline that creates differences among 10 subjects.
This process generates surface files for cortical and white matter surfaces 
(see~\cite{fischl2012freesurfer} for more detail of the process). 
Also, FreeSurfer did not introduce any differences in 10 subjects.

Results show that the static builds of FreeSurfer in which uses the embedded libraries, 
significantly avoids the creation of differences.
However, we found that between-OS differences occur when pre-processed data are different in FreeSurfer. 
For this purpose, in the second test, the results of PreFreeSurfer feed to the FreeSurfer 
pipeline on each operating system separately.
Figure~\ref{fig:tissue_class} shows between-OS differences  
as a result of propagation and amplification of differences in FreeSurfer. % that coming from PreFreeSurfer outputs. 
This figure illustrates the sum of binarized differences 
from 20 subject results of the brain tissue segmentation across OSes.
The magnitude of differences is in the range of 1-20 as the total number of subjects.

\begin{figure}
%  \includegraphics{brain\_classification}
\centering
  \includegraphics[width=\columnwidth]{images/brain_classification.png} 
  \caption{Heatmap binarized differences between brain segmentation results (file \emph{aseg.hires.mgz}) from 
          20 subjects in FreeSurfer (CentOS6 vs. CentOS7).} 
  \label{fig:tissue_class}
\end{figure}

@@

Although we identify the processes in the pipeline that create differences, 
the amplification of these differences should be quantified in the future works.
For this purpose, we can evaluate the final result after perturbing the identified 
processes in the pipeline iteratively.
As a result, the processes that amplify differences would have a significant effect on the results.

>>>>

\usepackage{subcaption}
\usepackage{xspace}
\usepackage{stmaryrd} % for llbracket and rrbracket
\usepackage{amsmath}
\usepackage{ulem}
\usepackage{pifont}
\DeclareMathOperator*{\argmin}{argmin}

@@

Recently, validation of the reproducibility has been widely investigated 
in the field of neuroimaging in which uses optimized 
processing methods with the purpose of functional and structural 
assessments of the human brain.
According to the previous 
studies, the variety of computing infrastructures including workstation 
types, parallelization methods, operating systems, and analysis 
packages are known to influence the reproducibility of the analyses~\cite{Gronenschild2012, 
diethelm2012limits, Glatard2015, bowring2019exploring}.
These irreproducibility issues are reported as the result of the 
creation, propagation, and amplification of small numerical 
differences.

In particular, the effect of the operating systems on  
computational pipelines shows the creation of small numerical differences~\cite{Glatard2015, Scaria2017}.
These differences mainly correspond to the mathematical functions implemented 
in different operating system libraries.
For instance, changing the mathematical functions like \emph{expf()} and 
\emph{cosf()} which manipulate the precision of floating-point representations, 
between \emph{glibc} libraries in different operating systems can produce 
small numerical differences.
Besides, a similar issue is expected for any operating system which is 
based on \emph{glibc}, the GNU C library.

There are different approaches to improve the reproducibility of the analysis, 
including containerization techniques that encapsulate software/hardware dependencies, 
provenance capturing tools, and version control systems. 
However, a comprehensive solution requires to fix the numerical instabilities instead of 
masking the problem. For this purpose, bagging is one solution that has been used to 
stabilize motion estimation analysis in fMRI~\cite{Glatard2018hbm}. 
Bootstrap aggregation reduces the instabilities, but it needs expensive computations.
As the second solution, a debugging tool can help to identify and 
fix the instabilities in the pipeline. Therefore, we introduce 
the Numerical Reproducibility Measurement (NURM) tool to identify 
the processes in the pipeline that create numerical differences across different conditions. 
Furthermore, we can stabilize the identified processes in the next steps.

The remainder of this paper describes the NURM-tool in different parts 
including the way of capturing and representation of the provenance 
information, the method of clustering of different subject types, and 
process labelling to characterize differences. We describe the HCP 
pre-processing pipelines as a popular neuroimaging project to test the 
NURM-tool and evaluate its functionality. Finally, we reported the 
experimental results which describe the processes that are responsible 
for the differences in the pipelines.

@@

\begin{figure}
\centering
  \includegraphics[width=0.3\columnwidth]{images/simple_graph}
  \caption{Process tree and graph
  constructed from the example pipeline in
  Listing~\ref{listing:sample-script}.
  Processes that read or write
  temporary files are 
  represented with squares. Plain edges 
  represent the process tree (\texttt{fork()} or \texttt{clone()} 
  system calls). Dashed edges represent file dependencies: temporary 
  files are in purple and result files are in black.
  Every node in the tree is labeled using (1) a process id created by our
  reconstruction, (2) the name of the executable run by the process.
  Process 0 is the initial call to the \texttt{sh} interpreter, and
  processes 1, 3 and 4 are the calls to FSL bet, stats and maths made in
  Listing~\ref{listing:sample-script}. Process 2 is forked by process 1: it
  was captured by \reprozip while it did not appear in
  Listing~\ref{listing:sample-script}. 
}
  \label{fig:simple_script}
\end{figure}

@@

We label in two categories the processes of the graph of each subject. First,
processes that read files that do not have differences and write files that
do not have differences are labeled \emph{transparent}
(Figure~\ref{fig:green}). Second, processes that read files that do not
have differences but write files that have differences \emph{create}
differences (Figure~\ref{fig:red}). In addition, processes that read files
that have differences and write files that also have differences are called
\emph{undetermined} (Figure~\ref{fig:grey}). Processes are also labeled
undetermined when they read or write temporary files, or produce multi-write
files. To resolve undetermined processes, hooks must be added to the pipeline to 
execute these processes on files without differences, and to capture
intermediary file states.

\begin{figure}%\centering
\centering
    \begin{subfigure}{0.3\linewidth}
      \centering
        \includegraphics[scale=0.36]{images/green.png}
        \caption{Transparent}
        \label{fig:green}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.35\linewidth}
      \centering
    \includegraphics[scale=0.36]{images/red.png}
    \caption{Creates differences}
    \label{fig:red}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.3\linewidth}
      \centering
    \includegraphics[scale=0.36]{images/grey.png}
    \caption{Undetermined}
    \label{fig:grey}
\end{subfigure}
    \caption{Different type of labeled process based on the input/output files.
  Dashed edges refer to the file dependencies between processes A and B 
  if a file written by A is read by B.}
    \label{fig:processes}
\end{figure}

@@

  # Run the original executable using the same arguments
  executed_file = sys.argv[0]
  exec_proc = op.join('original_scripts', executed_file)
  exec_argv = " ".join(map(pipes.quote, sys.argv[1:]))
  exec_cmd = exec_proc + exec_argv
  subprocess.Popen(exec_cmd, shell=True).communicate()

@@

  if exec_cmd in process_list.keys():
      # Save output files to read-only directory
      p_files = process_list[exec_cmd]
      shutil.copy(p_files, '/read-only/dir')

@@

Since all the result files are captured in Condition 1, the tool 
is ready to label the piepeline processes based on the results in Condition 2.
However, another challenge is to label undetermined processes that read files with 
differences and write files with differences too. In this
situation, it is not possible to determine from the pipeline results 
whether the process created differences or 
was transparent.
To address this issue, we set hooks in the pipeline 
in Condition 2. Hooks are set by replacing the pipeline 
processes with a custom script. This custom script labels each process 
as \emph{transparent} or \emph{create differences} immediately after running 
the original executable. 
The replacement is done through the PATH variable, as before. Moreover, 
this custom script copies the results obtained from the same process in 
Condition 1 to the pipeline output in Condition 2 if 
it is invoked with the arguments of a process that created differences 
(Figure~\ref{fig:overview-labelling} - (3)). 
Differences are identified by comparing checksum of the output files for 
the same process in both conditions (Figure~\ref{fig:overview-labelling} - (2)).
In addition, the distance functions that are specific to file types can be added 
to our tool to measure differences more precisely. 
We developed an incremental approach that consists of the following steps:

@@

Furthermore, the average time per subject taken for PreFreeSurfer processing is $\approx$ 2.2 hours, 
and FreeSurfer processing is $\approx$ 8.3 hours (20 runs), 
and the average output file size of 20 subjects in PreFreeSurfer is $\approx$ 2.7 GB, 
and FreeSurfer is $\approx$ 4.1 GB.

HCP data collected from different types of imaging techniques, including 
structural imaging (sMRI), functional imaging (fMRI), and diffusion imaging (dMRI).
The structural images include T1-weighted (T1w) and T2-weighted (T2w) images which 
help in the diagnosis of the brain injury.
The images are indexed with a suffix if several scans of the same modality were acquired.
For example, some data may include two T1w images with low and high resolutions.
The functional images, including task-based fMRI and resting-state scans, 
enable the measurement of functional activations within brain areas. 
Diffusion imaging is another kind of MRI technique, which measures 
the anatomical connectivity between regions.
See~\cite{van2013wu} for a discussion of these various modalities.

Data from 1200 subjects 
is available in \href{https://db.humanconnectome.org}{the ConnectomDB repository}. 
In this paper, we randomly selected 20 unprocessed subjects 
containing the 3-tesla type of MRI scanners from the HCP data release S500.

@@

We didn't observe any within-OS difference in PreFreeSurfer, however, we
identified two FreeSurfer processes leading to within-OS differences due
to the use of pseudo-random numbers: image registration with
\emph{mri\_segreg}, and cortical surface curvature
estimations with \emph{mris\_curvature}. To address this issue, we fixed
the seed used by the pseudo-random number generator, through defining 
option ``--seed'' in the command-line of the processes.
Our analysis can focus on between-OS differences.

@@

\begin{tabular}{cccccc}
\toprule
       &                        &  \multicolumn{4}{c}{Available Modalities}    \\ 
\cmidrule(lr){3-6}       
Type   &   \makecell{Number of \\ Subjects}   &  T1w\_1          & T1w\_2      & T2w\_1          & T2w\_2        \\ \midrule
1      &               9                      &   \ding{51}      &   \ding{51} &   \ding{51}     &   \ding{51}   \\ 
2      &               8                      &   \ding{51}      &             &   \ding{51}     &               \\ 
3      &               1                      &   \ding{51}      &             &   \ding{51}     &   \ding{51}   \\ 
4      &               2                      &   \ding{51}      &   \ding{51} &   \ding{51}     &               \\ 
\bottomrule
\end{tabular}
\begin{tablenotes}
     \small
     \item *Type refers to the topology type of the process tree of subjects.
\end{tablenotes}
\end{threeparttable}
\end{table}

Figure~\ref{fig:pfs_freq} shows the frequency at which  processes create differences 
in the pipeline, and how differences are distributed among sub-pipelines.
Differences were observed in linear registration 
with FSL FLIRT (in ACPC-Alignment, Brain Extraction, Distortion Correction, 
Atlas Registration), in non-linear registration with FSL FNIRT (in Brain Extraction 
and Altas Registration), and in image warping with FSL \emph{new\_invwarp} (in Brain Extraction 
and Atlas Registration). Differences were also observed in image mean 
computations with FSL maths  (in Anatomical Average).

Results show the variability between processes when 
the same tool behaves differently in different process instances. For example, 
FLIRT tool always produces files without differences in \emph{AAve} sub-pipeline but 
mostly creates different files in \emph{AAli} sub-pipeline.

Furthermore, different frequencies show the variability between subjects.
Results show that the same process can be reproducible for some subjects but not for others. 
Between subject differences can be a result of different subject types. 
The similar findings in the other studies show that the subject 
is an important factor in reproducibility~\cite{bowring2019exploring}.

% make figure in two columns use {figure*}
\begin{figure}
\centering
  \includegraphics[width=\columnwidth,height=4cm]{images/pfs_heatmap.png}
  \caption{Heatmap frequency of processes along with the sub-pipelines that introduce 
           differences among 20 subjects. 
           Each column represents the pipeline tool, and each row shows the sub-pipeline.
           Each cell may contain several process instances of a tool in a sub-pipeline.
           Annotations describe the frequency distribution of the instances as ``N [M] x Y" 
           in which N is the percentage of the process 
           that creates differences out of all M times that is occurred in the sub-pipeline. 
           Also, times Y shows the number of instances with the same distribution.
           For example, the cell of FLIRT tool in sub-pipeline AAve shows 10 process instances  
           with the frequency of 10 that introduce no differences.
           The range between red and green colors shows the highest and lowest 
           proportion of occurrences of the processes with differences.
           Also, grey cells indicate that the pipeline tool is not executed in the sub-pipeline.}
  \label{fig:pfs_freq}
\end{figure}

Figure~\ref{fig:fnirt_result} 
shows the checkerboard images from the results of ACPC-Alignment in PreFreeSurfer. 
Each rectangle includes the replaced pixels with different intensity values between images.
The misaligned regions visually represent significant differences in the alignment 
procedure.

@@

\begin{figure}
  \centering
    \includegraphics[width=\columnwidth]{images/t2w_alignment.png} 
    \caption{Differences between FNIRT results of one subject from PreFreeSurfer 
    (checkerboard image \emph{T2w\_acpc\_to\_MNI\_nonli.nii.gz}) (CentOS6 vs. 
    CentOS7). Sagittal (top) and axial (bottom) views of the brain in dimension of X=130, Y=155, Z=130.
    Right images show the zoomed version part of the images marked by the red rectangle.
    Here is 
    \href{https://github.com/ali4006/HCP-reproducibility-paper/blob/master/images/pfs_t2w_alignment.gif}
    {the link} to the gif animation of the same images.
} 
    \label{fig:fnirt_result}
\end{figure}

@@

NURM-tool identified \emph{mris\_make\_surfaces} as the only process in 
the pipeline that creates differences in 10 subjects among 20 subjects.
This process generates surface files for cortical and white matter surfaces 
(see~\cite{fischl2012freesurfer} for more detail). 
Also, FreeSurfer did not introduce any differences in 10 subjects.

Results show that the static builds of FreeSurfer in which uses the embedded libraries 
significantly avoids the creation of differences.
However, we found that between-OS differences occur when pre-processed data are different in FreeSurfer. 
FreeSurfer may not introduce differences, but it propagates and amplifies 
the current differences that appeared in PreFreeSurfer results across the operating systems.

Figure~\ref{fig:tissue_class} illustrates the sum of binarized differences 
from 20 subject results of the brain tissue segmentation across OSes.
This figure shows between-OS differences  
as a result of the propagation and amplification of differences in FreeSurfer. 
Differences are mostly localized at the interfaces between tissues.

Furthermore, Figure~\ref{fig:scatter_plot} plots the mean of Dice coefficients 
for the 45 segmented regions obtained on two operating systems.
High Dice coefficients (more than 0.95) in most of the regions indicate minor differences.
Dice values can be very low , down to 0.5, in particular in small structures 
such as the vessels and lateral ventricles.

\begin{figure}
%  \includegraphics{brain\_classification}
\centering
  \includegraphics[width=\columnwidth]{images/brain_segmentation_mni.png} 
  \caption{Heatmap binarized differences between brain segmentation results (file \emph{aseg.hires.mgz}) from 
          20 subjects in FreeSurfer (CentOS6 vs. CentOS7). 
          All binarized differences were resampled to the default MNI152 volume template.
          Here is 
          \href{https://github.com/ali4006/HCP-reproducibility-paper/blob/master/images/fs_brain_segmentation.gif}
          {the link} to the gif animation from a single subject.} 
  \label{fig:tissue_class}
\end{figure}

\begin{figure}
  \centering
    \includegraphics[width=\columnwidth]{images/scatter_plot.png} 
    \caption{The scatter plot of the mean of Dice coefficients between segmented regions 
            from 20 subjetcs in FreeSurfer (CentOS6 vs. CentOS7). 
            The complete Dice coefficient values are 
            \href{https://github.com/ali4006/HCP-reproducibility-paper/blob/master/data/dice_values.csv}
            {available}.} 
    \label{fig:scatter_plot}
  \end{figure}

@@

Although we identify the processes in the pipeline that create differences, 
the amplification of these differences should be quantified in the future works.

\note{FREESURFER RESULTS: 
These are consistant results to the paper~\cite{Glatard2015}, as it is shown the propaggation of differences in cortical thickness extraction analysis.
The magnitude of differences from the operating system changes is out of our expectaion}
